---
title: "Confidence Intervals"
author: "Ben Levy"
date: "2022-08-01"
output: html_document
---

```{r}
library(moderndive)
library(tidyverse)
library(infer)
```

# Bootstrapping

EDA of `pennies_sample`

- n = 50 pennies
- Goal: estimate the average age of all US pennies (measured in 2019)

```{r}
pennies_sample %>%
    ggplot(aes(x = year)) +
    geom_histogram(fill = "chartreuse", color = "coral", binwidth = 3)
```

- We are assuming that this sample (n = 50) is representative of all US pennies

```{r}
summary(pennies_sample$year)
x_bar <- mean(pennies_sample$year)
```

- **Key**: Understanding the *variability* in our sample mean

## Bootstrapping Rules
    - Must be done with replacement sampling from our **original sample**
    - Bootstrapped sample is the same size as the **original sample**
    
```{r}
bootstrap_sample1 <- rep_sample_n(pennies_sample, size = 50, reps = 1, replace = TRUE)

bootstrap_sample1 %>%
    ggplot(aes(x = year)) +
    geom_histogram(binwidth = 3)

# bootstrap statistic
mean(bootstrap_sample1$year)
```

- If repeating this, the *bootstrap distribution* will be centered around the sample mean.

- The bootstrap procedure works as if our sample (i.e., many copies of our sample) were the population

## Bootstrapping with `infer`

```{r}
pennies_sample %>%
    specify(response = year) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "mean") %>%
    visualize()
```

## Confidence Intervals

Confidence level determines how wide the interval is.

Common intervals: 90%, **95%**, 98%, 99%

Using the percentile method to obtain confidence intervals

- Involves finding the *quantiles* for the middle 95% (or 90% or 99%) of the bootstrap distribution

```{r}
pennies_sample %>%
    specify(response = year) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "mean") %>%
    get_ci(level = 0.95) -> percentile_ci
```

We are 95% confidence that the population mean penny year is between 1991.2 and 1999.8.

```{r}
pennies_sample %>% 
  specify(response = year) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean") %>%
  visualize()  + 
  shade_ci(endpoints = percentile_ci, 
           color = "hotpink", fill = "chartreuse")
```

## Interpreting CIs

When we say we are "95% confident," what does that actually mean?

To answer this, let's calculate a confidence interval for a **proportion** rather than a *mean*.

- Treat the `bowl` dataset as our *population*.
    - This means we know what the population proportion is ($p_{red}$ = 0.375)
- Use the sample size of n = 90, and calculate a CI for the proportion of red

```{r}
bowl %>%
    slice_sample(n = 90) %>%
    specify(response = color, success = "red") %>% # need to specify a success option when calculating a proportion
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "prop") %>%
    get_ci(type = "percentile")
```

CI: (0.310, 0.511): this includes the true $p$.

If we were to repeat this procedure over and over again with different samples, approximately 95% of those samples would yield a 95% CI that covers the true parameter.

## Width of CI

Two ways to impact the width of CI:

- Change the confidence level
    - 99% is wider than 98%
- Sample size
    - Narrower CIs when increasing sample size
    
## Theory-Based CI using the CLT

- Relies on the conditions that our sample is large neough, the data come from a normal distribution, and that our sample was collected at random.

```{r}
mean(pennies_sample$year) - 1.96*(sd(pennies_sample$year)/sqrt(50))
mean(pennies_sample$year) + 1.96*(sd(pennies_sample$year)/sqrt(50))
```

