---
title: "Data Cleaning"
output: html_notebook
---

# Data Cleaning

```{r}
library(tidyverse)
library(fivethirtyeight)

spotify_songs <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv")
bush_approval_untidy <- read_csv("./bush_approval.csv")
```

## Tidy Data

```{r}
bush_approval <- bush_approval_untidy %>%
  select(date, Republican, Independent, Democrat)
```

Instead of this format, we want a dataset with the following three columns: date, party, approval

```{r}
bush_approval <- bush_approval %>%
  pivot_longer(!date, names_to = "party", values_to = "approval")
```

Visualizing Tidy Data:

```{r}
bush_approval %>%
  ggplot(aes(x = date, y = approval, color = party)) +
  geom_line(size = 20)
```

Visualizing Un-Tidy Data:

```{r}
bush_approval_untidy %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = Republican), color = "lightblue3", size = 1.5) +
  geom_line(aes(y = Democrat), color = "chartreuse", size = 1.5) +
  geom_line(aes(y = Independent), color = "lightpink2", size = 1.5)
```

## Data Cleaning Tips and Tricks

### `count()`

```{r}
video_games <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv")
```

#### Questions

- What developers appeared the most frequently?
  - Choice of Games
- How many developers are there?
  - 16,753 (number of rows in DF)

```{r}
video_games %>%
  count(developer, sort = TRUE)
```

Suppose we have a categorical variable with way too many levels to count or use (e.g., `publisher` or `developer`). We can use the function `fct_lump()` less frequently occurring levels in a categorical variable.

```{r}
video_games_count <- video_games %>%
  mutate(publisher_lump = fct_lump(publisher, n = 5)) %>%
  count(publisher_lump, sort = TRUE) %>%
  filter(publisher_lump != "Other") 
video_games_count %>%
  ggplot(aes(x = publisher_lump, y = n)) +
    geom_col(fill = "chartreuse") +
    coord_flip()
```

If we want to custom-order the levels of `publisher_lump`, use `fct_relevel()` within a `mutate()`.

```{r}
video_games_count %>%
  mutate(publisher_relevel = fct_relevel(publisher_lump, "SEGA", "Strategy First", "Big Fish Games", "Square Enix", "Ubisoft")) %>%
  ggplot(aes(x = publisher_relevel, y = n)) +
  geom_col() +
  coord_flip()
```

## `separate()`

This function separates one column into multiple columns.

```{r}
nyc_restaurants <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-12-11/nyc_restaurants.csv")
```

Split `inspection_type` into two columns.

```{r}
nyc_restaurants_sep <- nyc_restaurants %>%
  separate(col = inspection_type, into = c("inspection_program", "inspection_type"), sep = "/")

nyc_restaurants_sep %>%
  count(inspection_type, sort = TRUE)
```

Opposite of this is `unite()`.

## Filtering using Text Expressions

```{r}
nyc_restaurants %>%
  mutate(roach_violation = str_detect(string = violation_description, pattern = "roach")) -> nyc_restaurants

nyc_restaurants %>%
  filter(roach_violation = TRUE) %>%
  ggplot(aes(x = fct_infreq(boro))) +
  geom_bar()
```

