---
title: "dataWranglingMultipleTables"
author: "Ben Levy"
date: "2022-07-21"
output: html_document
---

# Data Wrangling with Multiple Tables

```{r}
library(tidyverse)
library(nycflights13)
```

## Combining Multiple Tables

### `inner_join()`

- Join two tables together based on a common key between the tables
- Only keeps rows that have a key match in either table (i.e., if the keys don't match, get outta here)
  - Thus, use this if you want to only keep rows that have matching keys (i.e., no missing column data after joining)

#### Random Subset of 10 Rows from the `flight` dataset

- Merge airport names and location information to the `flights_sample` dataset

```{r}
set.seed(12345)
flights_sample <- flights %>%
  slice_sample(n = 10) %>%
  select(month, dep_delay, dest, carrier)

# inner join
flights_sample %>%
  inner_join(airports, by = c("dest" = "faa")) -> flights_with_airports
```

#### Merge airport names and location information to the `flights` dataset

```{r}
flights %>%
  inner_join(airports, by = c("dest" = "faa")) -> flights_inner
```


Let's use `setdiff()` to see the differences between the `flight$dest` and `airports$faa` vectors.

```{r}
setdiff(flights$dest, airports$faa) # see what airports that are in flights$dest that aren't in airports$faa
```


### `full_join()`

- Merging regardless of key matches

```{r}
flights_full <- flights %>%
  full_join(airports, by = c("dest" = "faa"))

flights_full %>%
  na.omit() # removes any row with ANY NA values
```


### `left_join()` and `right_join()`

#### `left_join()`

- Join two tables together, keep ALL rows in the left-hand table, regardless of whether they have a key match in the right-hand table.

```{r}
flights_sample_left <- flights_sample %>%
  left_join(airports, by = c("dest" = "faa"))
```

#### `right_join()`

- Same as `left_join()`, just with the right (i.e., the right table persists, regardless of whether they have a key match in the left-hand table).

```{r}
flights_sample_right <- flights_sample %>%
  right_join(airports, by = c("dest" = "faa"))
```