---
title: "Data Visualization"
author: "Ben Levy"
date: "2022-07-19"
output: html_document
---

# Imports

```{r}
library(tidyverse)
library(nycflights13)
```

# Grammar of Graphics

## Three required components for a **valid** data graphic

-   **Data**
-   **Aesthetic attributes**
    -   Makes the graph axis and other appearance deviations from the default
    -   E.g., x-position; y-position; size, color, and shape of points
    -   Maps from the data (mapping from variables in data to aesthetic attributes)
-   **Geometric object** or **geom**
    -   I.e., type of graph (e.g., line graph)
    -   `geom_type` (e.g., `geom_point`)
    -   Can specify certain stylistic variables (e.g., alpha [opacity])

## Optional Components

-   Labels, themes, point transparency, annotations, coordinate system
-   Labels
    -   Add using `labs()`

    -   Add title (`title=`)

    -   Add subtitle (`subtitle=`)

# Building ggplots

## Graph of the Relationship Between `wt` and `mpg` in the `mtcars` dataset.

Negative association (non-linear). I.e., as one increases, another decreases

```{r}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) + 
  geom_smooth() + 
  geom_point()
  
```

## 5 Common Plots

### Scatterplots

-   Effective for visualizing two **NUMERICAL** variables
-   Can add on other variables (e.g., a variable to color code the points)

Looking at the `flights` dataset (**correlation between flights' departure delay and arrival delay**), looking at only `F9` (Frontier):

```{r, warning = FALSE, message = FALSE}
nycflights13::flights %>%
  filter(carrier == "F9") %>%
  ggplot(mapping = aes(x = dep_delay, y = arr_delay)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", color = "red", size = 1.5)
```

### Linegraphs

-   Similar to scatterplots, but the `x` variable is usually **TIME-RELATED** (a continuous variable).

A great example is weather from `JFK` from November 1 through November 18 by the `datetime`:

```{r}
nycflights13::weather %>%
  filter(origin == "JFK", month == 11 & day <= 18) %>%
  ggplot(mapping = aes(x = time_hour, y = temp)) +
  geom_line(size = 1, color = "blue") +
  geom_point(alpha = 0.25, size = 2) +
  labs(x = "Date", y = "Temperature (F°)", title = "Hourly Temperature Recordings from JFK Airport", subtitle = "November 1-18 in 2013") +
  theme_linedraw() +
  theme(plot.title = element_text(color = "blue", hjust = 0.5, face = "bold"), plot.subtitle = element_text(hjust = 0.5))
```

### Histograms

-   Graphics for visualizing the **distribution** of a **SINGLE** **numerical** variable
-   You can **ONLY** specify the x-axis label
-   Items to look at in histograms
    -   Center (i.e., median)
    -   Shape (i.e., skewedness)
    -   Spread (i.e., range of data, range of certain percentile)
- **NOT A BAR GRAPH**
  - As histograms can only contain **numerical** data
- The `fill=` parameter can be used to change the background color of each bar
- Facets
  - Split a dataset by something (e.g., month)

Looking at the temperatures in the `nycflights13::weather` dataset by count

```{r, warning = FALSE, message = FALSE}
nycflights13::weather %>%
  ggplot(mapping = aes(x = temp)) +
  geom_histogram(color = "white", fill = "hotpink") +
  facet_wrap(~month) +
  labs(x = "Temperature (F°)", title = "Temperatures by Month in Three Main Airports in NYC") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

-   The shape of this is **bimodal** (i.e., two main peaks)
    -   There are two main "modes"

### Boxplots

- Boxplots are similar to histograms, but they are **NOT** the same
- Boxplots show the distribution of a single numerical variable (same as histograms)
  - Side-by-side boxplots with a **categorical x** and **numerical y**
- Boxplots visualize **five** points
  - Minimum
  - First quartile (25th percentile: the first 25% of the data)
  - Median
  - Third quartile (75th percentile: the first 75% of the data)
  - Maximum
- The box represents the middle 50% of an ordered dataset (interquartile range: IQR)

```{r}
summary((nycflights13::weather %>%
  filter(origin == "JFK", month == 11 & day <= 18))$temp)
```

Visualizing the relationship between month and temperature using side-by-side boxplots using `nycflights13::weather`:

Note: *We need to force `month` to be categorical in order for the boxplots to display side-by-side.* Can do this using `factor(x)`, where `x` is the x-variable.

```{r, warning = FALSE}
nycflights13::weather %>%
  ggplot(mapping = aes(x = factor(month), y = temp)) +
  geom_boxplot() +
  labs(x = "month")
```

### Barplots/Bar graphs

- Categorical x-value(s)
- y-axis shows count (or other summary statistic [e.g., mean])
- Pre-counted data: `geom_bar()` - specify x-aesthetic
- Counted data: `geom_col()` - specify both x- and y-aesthetic

Making a bar graph of the pre-counted data: `carrier` from `nycflights13::flights`:

```{r}
nycflights13::flights %>%
  ggplot(mapping = aes(x = carrier)) +
  geom_bar() +
  coord_flip()
```

Finding the relationship between carrier and origin (airport):

- Map fill to the `origin`
  - Adjusts the colors of portions of bars to certain carriers
- Adding `position = "fill"` to the parameters of `geom_bar(...)` normalizes the bar graphs (allowing for better comprehension of the graph)

```{r}
nycflights13::flights %>%
  ggplot(mapping = aes(x = carrier, fill = origin)) +
  geom_bar(position = "fill") +
  coord_flip()
```
