---
title: "Data Wrangling"
author: "Ben Levy"
date: "2022-07-20"
output: html_document
---

# Data Wrangling

```{r}
library(tidyverse)
library(nycflights13)
library(gapminder)
```

## (Non-Native) Pipe Operator: `%>%`

-   Behaves just like a literal pipe (`|`) in Unix-based systems

### Filtering All Flights Carried by JetBlue & Summarize the Flight Distances

```{r}
nycflights13::flights %>%
  filter(carrier == "B6") %>%
  group_by(origin) %>%
  summarize(mean_distance = mean(distance))
# analogous to:
summarize(group_by(filter(nycflights13::flights, carrier == "B6"), origin), mean_distance = mean(distance))
```

## Five Main Function for Data Wrangling with DPlyer (D-Plier)

-   `select()`
    -   Subset of columns (e.g., SQL: `SELECT <column(s)> FROM <table>`)
-   `filter()`
    -   Subset of rows (e.g., SQL: `SELECT * FROM <table> WHERE <condition(s)> = <value(s)>`)
-   `mutate()`
    -   Add or modify existing columns
-   `arrange()`
    -   Sort rows
-   `summarize()`
    -   Aggregate data in some shape

### `select()`

```{r}
gapminder %>%
  select(country, year, lifeExp) -> gapminder_condensed
```

### `filter()`

#### Get all rows for the United States

```{r}
gapminder %>%
  filter(country == "United States") -> gapminder_us
```

#### Get all rows for the United States or Canada from 1980-present

```{r, warning=FALSE, message=FALSE}
gapminder %>%
  filter((country == "United States" | country == "Canada") & year >= 1980) -> gapminder_combo
#OR
gapminder %>%
  filter(country %in% c("United States", "Canada") & year >= 1980) -> gapminder_combo_in 
  # %in% checks if the country column's value is within the following vector
gapminder_combo_in %>%
  select(year, country, lifeExp) %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) +
  geom_point() +
  geom_smooth()
```

### `arrange()`

#### Sort by year and continent in ascending order

```{r}
gapminder %>%
  arrange(year, continent) 
  # specifying more than one parameter (column to sort by) will simply give the first priority over the second, etc.
```

#### Sort by descending order (wrap with `desc()`)

```{r}
gapminder %>%
  arrange(desc(pop))
```

#### Extract the top 5 recordings in terms of `pop`

- Use `slice_max()`
  - `slice_min()` starts from the minimum

```{r}
gapminder %>%
  filter(year < 2000) %>%
  slice_max(pop, n = 5)
```

### `mutate()`

- Adds columns to an existing dataset

#### Convert degrees F to degrees C in the `nycflights13::weather` dataset

```{r}
nycflights13::weather %>%
  mutate(temp_C = (temp-32)/1.8) -> weather
```

### `summarize()` with `group_by()`

- Another method to calculate basic stats

#### Calculating the mean and SD of temperatures from the `weather` dataset

- Use `na.rm = TRUE` in R stat functions to remove any null values

```{r}
weather %>%
  summarize(mean_F = mean(temp, na.rm = TRUE), sd_F = sd(temp, na.rm = TRUE))
```

#### Calculating these stats for each airport

- `group_by(...)` makes categories based on the parameter (a column)

```{r}
weather %>%
  group_by(origin) %>%
  summarize(mean(temp, na.rm = TRUE), sd(temp, na.rm = TRUE))
```

