---
 title: "Hypothesis Testing Part 1"
author: "Ben Levy"
date: "2022-08-02"
output: html_document
---

```{r}
library(tidyverse)
library(infer)
library(nycflights13)
movies_sample <- read_csv("./movies_sample.csv")
```

# When inference is not needed

## Difference in air time between NYC->BOS and NYC->SFO

```{r}
bos_sfo <- flights %>%
    na.omit() %>%
    filter(dest %in% c("BOS", "SFO"))
```

```{r}
bos_sfo %>%
    group_by(dest) %>%
    summarize(mean_air = mean(air_time), median_air = median(air_time), sd_air = sd(air_time))
```

On average, a flight NYC->SFO lasts about 300 minutes longer than NYC->BOS.

### Boxplots

```{r}
bos_sfo %>%
    ggplot(aes(x = dest, y = air_time)) +
    geom_boxplot()
```

### Density Plots

- x is numerical
- fill, if necessary, is categorical

```{r}
bos_sfo %>%
    ggplot(aes(x = air_time, fill = dest)) +
    geom_density()
```

$\mu_b = $ mean air time of all NYC->BOS
$\mu_s = $ mean air time of all NYC->SFO

$H_0$ -> $\mu_b$ - $\mu_s% = 0$ <-> $\mu_b = \mu_s$ 
$\mu_b-\mu_s != 0$ <-> $\mu_b != \mu_s$

# Basics of Hypothesis Testing

- These are always statements about the unknown population parameters

1. $H_0$ -> null hypothesis
    - Usually the no correlation hypothesis
2. $H_a$ -> alternative hypothesis
    - The claim we're trying to gather evidence for

- These claims are mutually exclusive and competing

3. Sample data - our data serve as our "evidence," which is, hopefully, in favor of the alternative.

**KEY**: start out by assuming the null is true

**Two possible conclusion**

1. Accept alt/Reject null
2. Fail to reject null - not enough evidence to accept alt

If we end up rejecting the null (accepting alt), our results are statistically significant

## **Example** Sarah the Chimp

Let p be the population proportion of correct decisions made by the chimp

- $H_0$: p = 0.5 (Sarah was just guessing)
- $H_a$: p > 0.5 (Sarah was doing better than guessing)

Our evidence, in the form of sample data, was the fact that Sarah made the correct decision *seven out of eight times*.
    - $\hat{p}$ = 7/8
    
Based on the mode of this experiment (if guessing) is 4/8. 7/8 is extremely unlikely to occur.

- This means that, maybe our "hypothetical" world where the null is true *is actually false*, and reality is more in line with the alternative.

# Hypothesis Testing in R

Using `infer`, there are five steps to follow, all arranged in a *pipeline*.

We will compare different movie genres' ratings

Question: what is the relationship between IMDB.com rating and genre (action vs romance) for movies?

$H_0$: $\mu_r$ - $\mu_a$ = 0: there is no association between genre and rating
$H_1$: $\mu_r$ = $\mu_a$ != 0: there is an association between genre and rating

## EDA

```{r}
movies_sample %>%
    group_by(genre) %>%
    summarize(mean_rating = mean(rating),
              median_rating = median(rating),
              sd_rating = sd(rating))
```

*sample diff in means*:

- $\bar{x}_r - \bar{x}_a = 1.15$; the question is *is our test statistic far enough away from the null value (0) to reject the null?*

```{r}
movies_sample %>%
    ggplot(aes(x = genre, y = rating)) +
    geom_boxplot()
```

The `infer` pipeline steps:

1. `specify()`
2. `hypothesize()`: specify the *type* of null hypothesis
    - `null = "independence"` if comparing *two groups*: genre and rating are independent
    - `null = "point"` if comparing a single group to a number
3. `generate()`: generate 1,000 resamples from a world where the null is true
4. `calculate()`: calculate statistics per sample from the *generate* step

In the generate step, we're simulating a world where the null is true.

- If we're assuming *no relationship between genre and rating*, then we can simulate the null hypothesis by shuffling all the ratings at random between genres.

```{r}
set.seed(34)

movies_sample %>%
    group_by(genre) %>%
    slice_sample(n = 10) -> movies_sample_smaller
```

```{r}
movies_sample %>%
    specify(response = rating, explanatory = genre) %>%
    hypothesize(null = "independence") %>%
    generate(reps = 10000000, type = "permute") %>%
    calculate(stat = "diff in means") -> null_distribution

null_distribution %>%
    get_p_value(obs_stat = 1.15, direction = "both")

visualize(null_distribution)+ # represents the diff in means from a world where the null is true
    shade_p_value(obs_stat = 1.15, direction = "both")
```

p-value: the probability of observing a sample statistic at least as extreme as ours in a world where the null is true; the percent of data exceeding the observed statistic, under the null hypothesis

**Our p-value is < 0.001: we say this because it is possible, given more reps, to have an outlier p-value.**

With this p-value, we must form a decision (i.e., reject or fail to reject the null hyp) and a conclusion.

- We compare the p-value to a pre-set *significance level*, denoted by $\alpha$. This is usually 5% (0.05).

1. If p < 0.05, reject the null and conclude the alternative
2. If p > 0.05, fail to reject the null

In our "study", p < 0.001, so we reject the null. Thus, there is a correlation between movie genre and rating.

## Relationship between sample size and p-value

```{r}
set.seed(2)

movies_smaller <- movies_sample %>%
    group_by(genre) %>%
    slice_sample(n = 10) %>%
    ungroup()

movies_smaller %>%
    group_by(genre) %>%
    summarize(mean_rating = mean(rating))

# diff of 2.39 (mu_r-mu_a)
```

