---
title: "hypothesisTesting2"
author: "Ben Levy"
date: "2022-08-03"
output: html_document
---

```{r}
library(tidyverse)
library(infer)
library(moderndive)
movies_sample <- read_csv("./movies_sample.csv")
```

# Day Two

$H_0: \mu = \mu_0$
$H_a: \mu != \mu_0$

## Point Hypotheses

Can we conclude, using our sample, that the population mean movie rating on IMDB is *different* from 6.38?

- $H_0: \mu = 6.38$
- $H_a: \mu < 6.38$ because movies are bad

### EDA

```{r}
movies_sample %>%
    ggplot(aes(x = rating)) +
    geom_histogram(color = "white", fill = "hotpink1")
```

Descriptive Statistics

```{r}
mean(movies_sample$rating)
```

### Analysis

```{r}
movies_sample %>%
    specify(response = rating) %>%
    hypothesize(null = "point", mu = 6.38) %>%
    generate(reps = 1000) %>% # shift center of distribution (mode) to 6.38 (the mu provided)
    calculate(stat = "mean") %>%
    get_p_value(direction = "less", obs_stat = 5.61)
```

p < 0.001; reject the null. We have enough evidence to conclude that the average IMDB movie rating is less than the 6.38 Some Guy claimed on Quora.

## Test for Comparing Two Proportions

We'll use `mythbusters_yawn` dataset to compare two proportions.

The Mythbusters wanted to test whether *yawning is contagious*.

- 50 participants recruited
- Each participant was "interviewed" by a member of the show
    - Group 1: interviewed by a "seed" who did yawn
    - Group 2: interviewed by a "control" who didn't yawn

$H_0:  p_s = p_c$
$H_a: p_s > p_c$

```{r}
mythbusters_yawn
```

### EDA

Because we have *two categorical variables*, let's look at stacked bar graphs

```{r}
mythbusters_yawn %>%
    ggplot(aes(x = group, fill = yawn)) +
    geom_bar(position = "fill")
```

Calculate sample diff in props:

```{r}
mythbusters_yawn %>%
    count(group, yawn)
```

```{r}
mythbusters_yawn %>%
    specify(response = yawn, explanatory = group, success = "yes") %>%
    calculate(stat = "diff in props")
```

Test:

```{r}
mythbusters_yawn %>%
    specify(response = yawn, explanatory = group, success = "yes") %>%
    hypothesize(null = "independence") %>%
    generate(reps = 1000) %>%
    calculate(stat = "diff in props", order = c("seed", "control")) %>%
    get_p_value(direction = "both", obs_stat = 0.04411765)
```

- If the null were true and yawning was not contagious, there would be ~50% chance of observing data at least as extreme as ours.

# Connecting Back to Confidence Intervals

Recap: confidence intervals provide a range of plausible values for the unknown population parameter.

If you use an infer pipeline for a hypothesis test, you can tweak it to get a CI by removing the hypothesize() step and change get_p_value to get_ci

Because 0 is in the CI (in Anthony's notes), we're claiming with 95% confidence that ZERO is a plausible value for $p_s - p_c$.
    - We cannot accept nor reject the null based on this because it includes 0 and other non-zero values
    
# Inference for Regression

```{r}
lm_fit <- lm(price ~ bedrooms + sqft_living + waterfront + sqft_living*waterfront, data = house_prices) 
get_regression_table(lm_fit)
```

