---
title: "multipleLinearRegressions"
author: "Ben Levy"
date: "2022-07-26"
output: html_document
---

```{r}
library(tidyverse)
library(moderndive)

house_prices <- house_prices %>%
  filter(bedrooms < 33)
```

# Multiple Regression

New formula: $\hat{y} = b_0 + b_1x_1 + ... + b_nx_n$

$\hat{price} = b_0 + b_1*bedrooms + b_2*waterfrontTRUE$

- The `b` coefficients are all **estimated** from DATA
- Does this association change when accounting for whether the home has a waterfront view?

### Requirements

At least:

- Numerical `y` (dependent)
- At ***least*** **TWO** `x` (explanatory) variables
  - Type doesn't matter (i.e., numerical or categorical)

### EDA

```{r}
house_prices %>%
  ggplot(aes(x = bedrooms, y = price, color = waterfront)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### Interaction Model

When we perform EDA (e.g., scatterplot), and it seems that the slopes are different, we might consider including an interaction term in the model.

*Note: adding a crossed feature (i.e., interaction term) is not always necessary*

```{r}
lm_price_interaction <- lm(price ~ bedrooms + waterfront + bedrooms*waterfront, data = house_prices)

get_regression_table(lm_price_interaction)
```

Equation from this LM: $\hat{price} = 114143.5 + 123862.3*bedrooms - 236296.0*waterfrontTRUE + 416652.0*(bedrooms*waterfrontTRUE)$

- The first two terms/coefficients correspond to the intercept and slope for **non-waterfront** homes
- The latter two correspond to the **difference in intercept** and **difference in slope**, respectively, when comparing non-waterfront homes to waterfront homes.
  - Offsets from non-waterfront homes
- The interaction term tells us that the slope for *waterfront* homes is actually more than *triple* the slope for non-waterfront homes (123862 versus 123862 + 416652)

### Parallel Slope Models

#### Assumptions

- The slopes corresponding to any categorical `x` variable are **parallel**
- Basically, just a multiple regression model *without* the interaction term

#### Visualizing using `geom_parallel_slopes()`

```{r}
house_prices %>%
  ggplot(aes(x = bedrooms, y = price, color = waterfront)) +
  geom_point() +
  geom_parallel_slopes()
```

```{r}
lm_price_parallel <- lm(price ~ bedrooms + waterfront, data = house_prices)

get_regression_table(lm_price_interaction)
```

- The coefficient of `bedrooms`
  - While keeping `waterfrontTRUE` constant, each additional bedroom is associated with a $`128265.22` increase in house price, on average.
  
#### Example using Penguin Data

```{r}
penguins = read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')
```

```{r}
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Two (or more) Numerical X Variables

Instead of considering `waterfront` as an X variable, let's model house price as a function of

1. Number of bedrooms per home
2. size of the home (`sqft_living`)

```{r}
house_prices %>%
  ggplot(aes(x = sqft_living, y = price)) +
  geom_point() +
  geom_smooth(method = "lm")
```

- For larger homes, the error (or loss; $y-\hat{y}$) increases
- Caution: variance in residuals is not constant; it increases as `x` increases. Exercise caution when using a linear regression model when we observe this.

```{r}
lm_multiple <- lm(price ~ bedrooms + sqft_living, data = house_prices)
get_regression_table(lm_multiple)
```

*Note: it may be tempting to conclude that there is now a negative association between number of bedrooms and price, but this is merely a result of accounting for the `sqft_living` variable*

Interpretation of the coefficient for bedrooms:

- After holding the size of the home fixed, for each addition bedroom, there is an associated decrease in the price of a home by about $62k.
  - Makes sense because as a home has more and more bedrooms, they get smaller, given the `sqft_living` is constant
  
## Model Selection

Going to use the `MA_schools` dataset from the `moderndive` package to model average math SAT score as a function of several variables

Model of SAT math score: `perc_disadvan` and `size`

### Task

- Perform EDA
  - Graphs/stats
- Build an *interaction* and *parallel slopes model*

```{r}
MA_schools %>%
  ggplot(aes(x = perc_disadvan, y = average_sat_math, color = size)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
interaction_model <- lm(average_sat_math ~ perc_disadvan + perc_disadvan*size, data = MA_schools)
get_regression_table(interaction_model)
```

The two interaction terms do not add much to the original slope for small schools.

```{r}
parellel_model <- lm(average_sat_math ~ perc_disadvan + size, data = MA_schools)
get_regression_table(parallel_model)
```

```{r}
schools_simple_model <- lm(average_sat_math ~ perc_disadvan, data = MA_schools)
get_regression_table(schools_simple_model)
```

The slope for `perc_disadvan` barely changes; a parallel slopes model is not super different from a simple linear regression.