---
title: "samplingDistributions"
author: "Ben Levy"
date: "2022-07-28"
output: html_document
---

# Sampling Distributions

$\hat{p} = sample proportion$
  - Estimates $p$, the population proportion
  - Used for categorical data

$\bar{x} = sample mean$
  - Used for numerical data
  - Estimates $M$, the population mean
  
```{r}
library(tidyverse)
library(moderndive)
```

# We will use *computer simulations* to *virtually sample* from a *virtual population*.

- Our *virtual population* will be the `bowl` dataset from moderndive.

```{r}
bowl
```

Our "population" consists of 2400 balls, which are either *either* red or white.

**Goal**: **Estimate** the population proportion of red balls ($\hat{p}$)

```{r}
virtual_sample <- rep_sample_n(bowl, size = 50)
```

Calculate sample proportion of red balls

```{r}
virtual_sample %>%
  summarize(prop = sum(color == "red")/n())
```

Calculate actual proportion of red balls ($p$)

```{r}
bowl %>%
  summarize(prop = sum(color == "red")/n())
```

Take 30 samples of size n = 50 from the `bowl` dataset:

```{r}
virtual_samples_30 <- rep_sample_n(bowl, size = 50, reps = 30)

virtual_prop_red <- virtual_samples_30 %>%
  group_by(replicate) %>%
  summarize(prop = sum(color == "red")/n())
```

```{r}
virtual_prop_red %>%
  ggplot(aes(x = prop)) +
  geom_histogram(binwidth = 0.05, color = "chartreuse")
```

Now, let's take 1,000 samples instead of 30!

```{r}
virtual_samples_1000 <- rep_sample_n(bowl, size = 50, reps = 1000)

virtual_prop_red_1000 <- virtual_samples_1000 %>%
  group_by(replicate) %>%
  summarize(prop = sum(color == "red")/n())

virtual_prop_red_1000 %>%
  ggplot(aes(x = prop)) +
  geom_histogram(binwidth = 0.05, color = "chartreuse")
```

So far, we've only been tweaking the *rep_count*.

```{r}
virtual_samples_100 <- rep_sample_n(bowl, size = 100, reps = 1000)

virtual_prop_red_100 <- virtual_samples_100 %>%
  group_by(replicate) %>%
  summarize(prop = sum(color == "red")/n())

virtual_prop_red_100 %>%
  ggplot(aes(x = prop)) +
  geom_histogram(binwidth = 0.05, color = "chartreuse")
```

- A sampling distribution is a distribution of sample statistics

Let's compare the sampling distributions we constructed with sample sizes of n = 50 and n = 100.

*Note: when we increased the sample size from 50 to 100, the shape and center stayed roughly the same (symmetric, center ~0.4). The spread DID decreased largely when increasing the sample size per rep*

```{r}
virtual_samples_10 <- rep_sample_n(bowl, size = 10, reps = 1000)

virtual_prop_red_10 <- virtual_samples_10 %>%
  group_by(replicate) %>%
  summarize(prop = sum(color == "red")/n())
```

```{r}
virtual_prop_red_10 %>%
  ggplot(aes(x = prop)) +
  geom_histogram(binwidth = 0.1, color = "chartreuse")
```

## Central Limit Theroem (for $\hat{p}$ )

- If sample is independent and sufficiently large, then $\hat{p} \dot\sim Normal(p, \sqrt{\frac{p(1-p)}{n}})$
  - The sampling distribution of $\hat{p}$ is approximately normal 
  - Tidle with dot under it is "approximately distributed as"
  - $p$ = mean of the population
  - SD = $\sqrt{\frac{p(1-p)}{n}}$
  
Let's see how well our simulations followed the CLT

```{r}
mean(virtual_prop_red_10$prop)
sd(virtual_prop_red_10$prop)

# CLT says the mean *should* be 0.375 (p)
sqrt(0.375*(1-0.375)/10)
```

### CLT for Means

Now, let's perform simulation studies for estimated the *pop mean* in stead of the *pop proportion*

Consider `pennies` as our population of 800 pennies.

Goal: use sample data to *estimate* the population mean `age_in_2011`.

```{r}
virtual_pennies_10 <- rep_sample_n(pennies, size = 50, reps = 1000)
virtual_pennies_10 <- virtual_pennies_10 %>%
  group_by(replicate) %>%
  summarize(mean_age = mean(age_in_2011))

virtual_pennies_10 %>%
  ggplot(aes(x = mean_age)) +
  geom_histogram(color = "white", binwidth = 2)
```

True mean:

```{r}
mean(pennies$age_in_2011)
```

## Central Limit Theroem (for $\bar{x}$ )

- If sample is independent and sufficiently large, then $\bar{x} \dot\sim Normal(\mu, \frac{\sigma}{\sqrt{n}})$
  - The sampling distribution of $\hat{p}$ is approximately normal 
  - Tidle with dot under it is "approximately distributed as"
  - $p$ = mean of the population
  - SD = $\sqrt{\frac{p(1-p)}{n}}$
  
